---
title: "Bayesian linear regression"
author: Dr Merlise A Clyde, Duke University
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Read in the data ###
```{r}
library(BAS)
data(bodyfat)
summary(bodyfat)
```

### Scatterplot and OLS line ###
```{r}
myblue = rgb(86,155,189, name="myblue", max=256)
mydarkgrey = rgb(.5,.5,.5, name="mydarkgrey", max=1)

bodyfat.lm = lm(Bodyfat ~ Abdomen, data=bodyfat)

plot(Bodyfat ~ Abdomen, data=bodyfat, 
     xlab="abdomen circumference (cm)", 
     col=myblue, pch=16, main="")
beta = coef(bodyfat.lm)
abline(beta, lwd=4, col=1)
```

### Posterior distribution using the reference prior###

We will use `lm` to obtain the OLS estimates which provide the posterior mean and standard deviation, while the `confint` function provides confidence intervals, which under the reference prior are 95% credible intervals.  The code below extracts them and just relabels the output.


```{r}
summary(bodyfat.lm)

out = summary(bodyfat.lm)$coef[, 1:2]
out = cbind(out, confint(bodyfat.lm))
colnames(out) = c("posterior mean", "sd", "2.5", "97.5")
round(out, 2)
```

### credible intervals for the mean and prediction ###

```{r}
x = bodyfat$Abdomen
y= bodyfat$Bodyfat
xnew <- seq(min(x), max(x), length.out = 100)
ynew <- data.frame(predict(bodyfat.lm, newdata = data.frame(Abdomen = xnew), 
                   interval = "confidence", level = 0.95))
plot(x,y, xlab = "abdomen", ylab="bodyfat", col=myblue, pch=16)
lines(ynew$lwr ~ xnew, lty = 2, lwd=3, col=mydarkgrey)
lines(ynew$upr ~ xnew, lty = 2, lwd=3, col=mydarkgrey)
abline(bodyfat.lm, col="orange")
ynew <- data.frame(predict(bodyfat.lm, newdata = data.frame(Abdomen = xnew), 
                   interval = "prediction", level = 0.95))
lines(ynew$lwr ~ xnew, lty = 3, lwd=3, col=mydarkgrey)
lines(ynew$upr ~ xnew, lty = 3, lwd=3, col=mydarkgrey)
points(bodyfat[39,"Abdomen"], bodyfat[39,"Bodyfat"], col="orange", cex=5)
legend(110,15, legend=c("Posterior mean", "95% CI for mean", "95% CI for predictions"), 
       col=c("orange",rep(mydarkgrey, 2)), lwd=3, lty=c(1,2, 3))

```

###Outliers###
```{r}
source("bayes-outliers.R")
library(mvtnorm)
outliers= Bayes.outlier.prob(bodyfat.lm)
prob.39 = outliers$prob.outlier[39]
prob.39

n = nrow(bodyfat)
k = qnorm(.5 + .5*.95^(1/n))
outliers.no= Bayes.outlier.prob(bodyfat.lm, k=k)
prob.no.39 = outliers.no$prob.outlier[39]
prob.no.39

```

