---
title: "Comparing two independent means"
author: Dr Merlise Clyde, Duke University
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

Define the data using summary statistics
```{r data}
y = c(52.1, 27.1)
sd=c(45.1, 26.4) 
n=c(22,22)
```

### Initialization ###

Source functions for slides.   There are a number of functions in this file to implement the methods (some of these have been cleaned up and added to the `statsr` package).  
```{r source}
source("behren-fisher.R")
myblue = rgb(86,155,189, name="myblue", max=256)
```

### MCMC Sampling ###
We are interested in the posterior distribution of $\mu_A - \mu_B$ under $H0$ (difference is exactly zero)  and under $H1$, the difference is not zero. 
We will run an MCMC sampler using R functions (this does not require `jags`).  This uses numerical integration using the R package `cubature` to obtain the marginal likelihoods that enter into the Bayes factors and then MCMC sampling to sample from the posterior distributions of parameters under H1.   Under H0 the difference is exactly 0 so no simulations are needed for $\mu_A - \mu_B$.  The `plot=T` option will produce a plot of the posterior distribution under H1 with the pointmass at 0 under H0 and with a shaded area corresponding to the 95% HPD region combining draws from H0 and H1.  *The hypotheses are labeled H0 and H1 here to match with this function output, while the video uses H1 and H2*.


```{r sim}
set.seed(5)
out = BFt.test(y = c(52.1, 27.1), grp=NULL, sd=c(45.1, 26.4), n=c(22,22),
               method="intrinsic",  suff=T,jags=F,nsim=10000,
               colHPD=myblue, main="", 
               xlab=expression(mu[A] - mu[B]), lwd=3, plot=T)
```

From the output we can extract the overall 95% HPD interval for the difference of the means and the posterior probability of H0 (that the means are equal).

```{r estimates}
out$ci
out$post.prob
```

This is based on simulating from the posterior distribution so the credible interval may change slighly if different random seeds are used or if more simulations are used. 



We can also compute the credible interval conditional on H1 that the means are different from the output using teh HPDinterval function from `coda`.

```{r ci-H2}
HPDinterval(as.mcmc(out$diff[out$diff != 0]))
```

Remember this is calculated assuming that the means are different and so does not include the "draws" from the posterior under H1 that the means are exactly zero.

The sample mean under H1 is 
```{r mean H2}
mean(out$diff[out$diff != 0])
```

The overall mean based on averaging over the two hypotheses is obtained by taking the mean under H0 times the posterior probabilty of H0 plus the posterior mean under H1 times the posterior  probabilty of H1:

```{r post mean}

out$post.prob.H0*0 + 
  (1 - out$post.prob.H0)*mean(out$diff[out$diff != 0])
```
Because there is substantial probability that the difference is zero this linear combination of the posterior means is shrunk towards the poeterior/prior mean under H0 of zero.

### Using JAGS ###

If you have installed `jags` on your computer, then the following option using `jags=TRUE` is much faster, although the estimates of the posterior probability are subject to Monte Carlo variation.  To reduce that you may need to use a larger number of simulations.  If you have not installed `jags` comment out the chunk below.

```{r jags}
set.seed(5)
#out = BFt.test(y = c(52.1, 27.1), sd=c(45.1, 26.4), n=c(22,22), 
#               method="intrinsic", colHPD=myblue, nsim=1000000,
#               suff=T, jags=TRUE, main="", 
#               xlab=expression(mu[A] - mu[B]), lwd=3, plot=T)

out.jags = BFt.test(y = y, grp=NULL, sd=sd, n=n,
               method="intrinsic",  suff=T,jags=TRUE, nsim=100000,
               colHPD=myblue, main="", 
               xlab=expression(mu[A] - mu[B]), lwd=3, plot=T)

```

Estimates from jags

```{r jags-estimates}
out.jags$ci
out.jags$post.prob
```

These will not agree completely with the estimates above as `jags` does not allow the limiting "improper" prior distributions that are used in the intrinsic prior.

```{r ci-H2-jags}
HPDinterval(as.mcmc(out.jags$diff[out.jags$diff != 0]))
```